---
title: "Measuring human rights in Europe"
output: html_document
---

## Exercises and Answers

1. **To perform good analysis, many times you have to get to know how a given indicator works. In these exercises we are going to use a data about PTS (the Political Terror Scale) and SVS (the Societal Violence Scale). Make a research what they are, how they are defined, when they are used, who uses them and why.**

 - [PoliticalTerrorScale](https://politicalterrorscale.org)
 - [PoliticalTerrorScale_codebook](https://politicalterrorscale.org/Data/Files/PTS-Codebook-V140.pdf)
 - [PoliticalTerrorScale_FAQ](https://politicalterrorscale.org/About/FAQ/)
 - [SocietalViolenceScale](https://www.politicalterrorscale.org/Data/Documentation-SVS.html)

2. **The differences between PTS scores assigned by different organizations are significant? What about difference between countries and years?** [dataset](https://www.politicalterrorscale.org/Data/Download.html)

Proposed answer:

```{r message = FALSE, error = FALSE, warning = FALSE, eval=T}
library(readxl)
library(dplyr)
library(tidyr)
library(FSA)
library(data.table)
PTS <- read_excel("PTS-2021.xlsx",
                  col_types = c("text", "text", "text", "text", "numeric",
                                "text", "numeric", "text", "numeric", "numeric",
                                "numeric", "numeric", "numeric","numeric"))
PTS <- na.omit(PTS)
PTS <- PTS[,c(1,3,9,10,11)]
PTS %>%  pivot_longer(
  cols = starts_with("PTS_"),
  names_to = "Organization",
  names_prefix = "PTS_",
  values_to = "score",
  values_drop_na = TRUE) -> PTS
```
```{r message = FALSE, error = FALSE, warning = FALSE, include=FALSE,eval=F}
library(readxl)
library(dplyr)
library(tidyr)
library(FSA)
library(data.table)
library(readr)
PTS <- read_csv("PTS")
PTS$Year <- as.factor(PTS$Year)
PTS$Country <- as.factor(PTS$Country)
PTS$Organization <- as.factor(PTS$Organization)
```
```{r message = FALSE, error = FALSE, warning = FALSE}
shapiro.test(lm(score ~ Organization, data = PTS)$residuals)

kruskal.test(score ~ Organization, data = PTS)

kruskal.test(score ~ Year, data = PTS)

kruskal.test(score ~ Country, data = PTS)

dunnTest(score ~ Country, data = PTS)$res %>% head()

PTS %>% mutate(CO=paste0(Country,"-",Organization)) -> PTS_2

kruskal.test(score ~ CO, data = PTS_2)

dunnTest(score ~ CO, data = PTS_2)$res %>% head()

results <- dunnTest(score ~ CO, data = PTS_2)$res

subset(results, P.adj<0.05)[,c(1,4)] %>% head()

kruskal.test(score ~ Country, data = subset(PTS,Organization=="H"))

subset(dunnTest(score ~ Country, data = subset(PTS,Organization=="H"))$res,P.adj<0.05)[,c(1,4)] %>% head()

kruskal.test(score ~ Country, data = subset(PTS,Organization=="S"))

subset(dunnTest(score ~ Country, data = subset(PTS,Organization=="S"))$res,P.adj<0.05)[,c(1,4)] %>% head()

kruskal.test(score ~ Country, data = subset(PTS,Organization=="A"))

subset(dunnTest(score ~ Country, data = subset(PTS,Organization=="A"))$res,P.adj<0.05)[,c(1,4)] %>% head()
```

3. **The difference between SVS scores between year is significant? Assume that $n*=n$ and $*=NA$, where $n$ is a SVG score.**

Proposed answer:

```{r message = FALSE, error = FALSE, warning = FALSE}
library(stringr)

SVS <- read_excel("SVS 2016 Aggregate.xlsx",
                  col_types = c("skip", "text", "text", "text", "text", "text"))

SVS %>% pivot_longer(
  cols = starts_with("20"),
  names_to = "year",
  values_to = "score",
  values_drop_na = TRUE) -> SVS

SVS$score <- as.numeric(str_remove(SVS$score, "\\*"))

SVS <- na.omit(SVS)

shapiro.test(lm(score ~ year, data = SVS)$residuals)

kruskal.test(score ~ year, data = SVS)
```

4. **Examine the relationship between old and new prices (dataset `homedata` from `UsingR` package). Old price is the independent variable. Predict the home price in the year 2000 that was worth 75 000 $ in 1970.**

Proposed answer:

```{r message = FALSE, error = FALSE, warning = FALSE}
library(UsingR)
?homedata

# linear rergression model
model <- lm(y2000 ~ y1970, data = homedata)

# prediction
predict(model, list(y1970 = 75000))
```

5. **Dataset `Salaries` from `carData` package contains the 2008-09 nine-month academic salary for Assistant Professors, Associate Professors and Professors in a college in the U.S. Create a regression model and see which variables affect the dependent variable `salary`. Predict `salary` value for your selected variable values.**

Proposed answer:

```{r message = FALSE, error = FALSE, warning = FALSE}
library(carData)
?Salaries

# multiple regression model
model <- lm(salary ~ ., data = Salaries)

summary(model)

# the best multiple regression model
model2 <- step(model)

newdata = data.frame(rank = "Prof",
                     discipline = "A",
                     yrs.since.phd = 30,
                     yrs.service = 35)

# prediction on newdata
predict(model2, newdata)
```